---
title: "Analyses"
author : "Jeremy Wicquart"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: "cosmo"
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 4
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# Load functions and packages

```{r}

# 1. Required packages ----

library(tidyverse) # Core tidyverse packages
library(readxl) # To read excel files
library(lubridate) # To deal with dates
library(broom) # To use models with tidyverse
library(RcppRoll) # For rolling functions
library(cowplot) # To assemble plots
library(vegan) # For NMDS
library(ggrepel) # To optimize text placement in ggplot

# 2. Import functions ----

source("functions/graphical_par.R")
source("functions/theme_graph.R")

# 3. Set theme_graph() as the default ggplot theme ----

theme_set(theme_graph())

# 4. Create figs directory if not existing ----

ifelse(!dir.exists("./../figs"), dir.create("./../figs"), FALSE)

```

# Load and modify data

## Bleaching prevalence

```{r}

# 1. Bleaching prevalence ----

# 1.1 File path --

file_path <- "./../data/raw/3. Gulf of Mannar bleaching intensity, 2005-2020.xlsx"

# 1.2 Import the first sheet --

data_bleaching <- read_xlsx(file_path, sheet = 2, col_types = "text") %>% 
  select(1:4) %>% 
  drop_na(LIT) %>% 
  rename(replicate = 1, island = 2, year = 3, bleaching = 4)

# 1.3 Loop to bind all the sheets --

for(i in 3:length(excel_sheets(file_path))){
  
  data_bleaching <- read_xlsx(file_path, sheet = i, col_types = "text") %>% 
    select(1:4) %>% 
    drop_na(LIT) %>% 
    rename(replicate = 1, island = 2, year = 3, bleaching = 4) %>% 
    bind_rows(data_bleaching, .)
  
}

# 1.4 Modifications --

data_bleaching <- data_bleaching %>% 
  mutate(year = as.numeric(as.character(year)),
         bleaching = as.numeric(as.character(bleaching)),
         date = as.Date(paste0(year, "-06-01")))

# 1.5 Calculate the mean and SD --

data_bleaching_mean <- data_bleaching %>% 
  group_by(island, year, date) %>% 
  summarise(bleaching_mean = mean(bleaching),
            bleaching_sd = sd(bleaching))

```

## Benthic community

```{r}

# 2. Benthic community ----

# 2.1 Import the different datasets --

# 2.1.1 Data from 2005 to 2018 --

data_benthic_2018 <- read_xlsx("./../data/raw/2. Benthic community structure data  (Year wise) between 2005 and 2018.xlsx", sheet = 14) %>% 
  select(-Region, -...21)

# 2.1.2 Data for 2019 --

data_benthic_2019 <- read_xlsx("./../data/raw/GoM data - 2019&2020.xlsx", sheet = 1) %>% 
  select(-Region)

# 2.1.3 Data for 2020 --

data_benthic_2020 <- read_xlsx("./../data/raw/GoM data - 2019&2020.xlsx", sheet = 2, n_max = 252) %>% 
  select(-Region)

# 2.2 Bind the datasets and misc. modifications --

data_benthic <- bind_rows(data_benthic_2018, data_benthic_2019, data_benthic_2020) %>% 
  select(-LCC) %>% 
  pivot_longer("ACB":"Abiotic", names_to = "category", values_to = "cover") %>% 
  drop_na(cover) %>% 
  filter(cover != 0) %>% 
  rename(replicate = LIT_Number, site = Site_Number, island = Island, year = Year) %>% 
  # Re-categorization
  mutate(subcategory = str_replace_all(category, c("ACB" = "Acroporidae Branching",
                                                   "ACT" = "Acroporidae Table",
                                                   "CM" = "Coral Massive",
                                                   "CB" = "Coral Branching",
                                                   "ACD" = "Acroporidae Digitate",
                                                   "ACF" = "Acroporidae Foliose",
                                                   "CF" = "Coral Foliose",
                                                   "CS" = "Coral Submassive",
                                                   "ACE" = "Acroporidae Encrusting",
                                                   "CE" = "Coral Encrusting")),
         category = ifelse(category %in% c("ACB", "ACT", "CM", "CB", "CF", "CE", "ACD", "ACF", "CS", "ACE"), 
                           "Hard corals", 
                           category),
         subcategory = ifelse(category %in% c("Algae", "CCA", "Others", "Abiotic", "Soft coral"), 
                              NA, 
                              subcategory),
         date = as.Date(paste0(year, "-06-01")))

# 2.3 Remove useless objects --

rm(data_benthic_2018, data_benthic_2019, data_benthic_2020)

```

## Coral recruits

```{r}

# 3.1 File path --

file_path <- "./../data/raw/6. recruitment, 2005 -2020.xlsx"

# 3.2 Import the first sheet --

data_recruits <- read_xlsx(file_path, sheet = 1, col_types = "text", skip = 1) %>%
  select(1:16) %>% 
  rename(genus = 1) %>% 
  drop_na(genus) %>% 
  pivot_longer(2:ncol(.), names_to = "year", values_to = "recruit") %>% 
  mutate(island = excel_sheets(path = file_path)[1])

# 3.3 Loop to bind all the sheets --

for(i in 2:length(excel_sheets(file_path))){
  
  data_recruits <- read_xlsx(file_path, sheet = i, col_types = "text", skip = 1) %>% 
    select(1:16) %>% 
    rename(genus = 1) %>% 
    drop_na(genus) %>% 
    pivot_longer(2:ncol(.), names_to = "year", values_to = "recruit") %>% 
    mutate(island = excel_sheets(path = file_path)[i]) %>% 
    bind_rows(data_recruits, .)
  
}

# 3.4 Modifications --

data_recruits <- data_recruits %>% 
  mutate(year = str_replace_all(year, c("2003-2005" = "2004",
                                        "Mar-2007" = "2007",
                                        "Mar-2008" = "2008",
                                        "Nov-2009" = "2009",
                                        "Nov-2010" = "2010",
                                        "Sep-2011" = "2011",
                                        "Nov-2012" = "2012")),
         year = as.numeric(year),
         recruit = as.numeric(recruit)) %>% 
  select(island, year, genus, recruit) %>% 
  mutate(island = str_extract(island, "[A-z]+"))

```

## SST data and DHW calculation

```{r}

# 4.1 Load data and misc. modifications --

data_sst_raw <- read.csv("./../data/02_sst-extracted.csv")

# 4.1.1 All islands --

data_sst_all <- data_sst_raw %>% 
  group_by(date) %>% 
  summarise(sst = mean(sst)) %>% 
  mutate(date = as.Date(date))

# 4.1.2 By islands --

data_sst_island <- data_sst_raw %>% 
  # Make the mean by island
  group_by(island, date) %>% 
  summarise(sst = mean(sst)) %>% 
  mutate(date = as.Date(date))

# 4.2 Calculate DHW --

# 4.2.1 All islands --

data_dhw_all <- data_sst_raw %>% 
  # Calculate mean sst by month for each site
  mutate(month = month(date)) %>% 
  group_by(month) %>% 
  summarise(mean = mean(sst)) %>% 
  ungroup() %>% 
  # Extract Maximum Monthly Mean (MMM) for each site
  filter(mean == max(mean)) %>% 
  # Calculate the bleaching threshold (MMM + 1째C)
  mutate(threshold = mean + 1) %>% 
  ungroup() %>% 
  # Join with the SST data
  bind_cols(data_sst_all, .) %>%
  # Calculate Degree Heating Weeks
  mutate(delta = ifelse(sst >= threshold, sst - threshold, 0)) %>% 
  arrange(date) %>% 
  mutate(dhw = roll_sum(x = delta, n = 84, align = "right", fill = NA)) %>% # 7 days * 12 weeks = 84 days
  select(-delta, -mean, -month) %>% 
  # Calculate mean across all year
  mutate(mean = mean(sst))

# 4.2.2 By islands --

data_dhw_island <- data_sst_raw %>% 
  # Calculate mean sst by month for each site
  mutate(month = month(date)) %>% 
  group_by(island, month) %>% 
  summarise(mean = mean(sst)) %>% 
  ungroup() %>% 
  # Extract Maximum Monthly Mean (MMM) for each site
  group_by(island) %>% 
  filter(mean == max(mean)) %>% 
  # Calculate the bleaching threshold (MMM + 1째C)
  mutate(threshold = mean + 1) %>% 
  ungroup() %>% 
  # Join with the SST data
  left_join(data_sst_island, .) %>%
  # Calculate Degree Heating Weeks
  group_by(island) %>% 
  mutate(delta = ifelse(sst >= threshold, sst - threshold, 0)) %>% 
  arrange(date) %>% 
  mutate(dhw = roll_sum(x = delta, n = 84, align = "right", fill = NA)) %>% # 7 days * 12 weeks = 84 days
  select(-delta, -mean, -month) %>% 
  # Calculate mean across all year
  mutate(mean = mean(sst))

# 4.3 Remove useless objects --

rm(data_sst_raw)

```

# Main figures

## Figure 1

```{r fig.height=9, fig.width=5}

# 1. Make the plots ----

# 1.1 SST plot --

plot_sst <- ggplot(data = data_dhw_all, aes(x = date, y = sst)) +
  geom_path(color = "#6c7a89") +
  geom_hline(yintercept = max(data_dhw_all$threshold), linetype = "dashed") +
  # DHW events
  geom_vline(xintercept = as.Date("2010-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2016-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2019-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2020-06-12"), col = "red") +
  # Labs and lims
  labs(x = NULL, y = "SST (째C)") +
  lims(x = c(as.Date("2005-01-01"), as.Date("2021-01-01"))) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# 1.2 DHW plot --

plot_dhw <- ggplot(data = data_dhw_all, aes(x = date, y = dhw)) +
  # Warning (NOAA)*
  annotate("rect", xmin = as.Date("2005-01-01"), xmax = as.Date("2021-01-01"), 
           ymin = 0, ymax = 4, fill = "#fcd670", alpha = .2) +
  # Alert level 1 (NOAA)*
  annotate("rect", xmin = as.Date("2005-01-01"), xmax = as.Date("2021-01-01"), 
           ymin = 4, ymax = 8, fill = "#eb9532", alpha = .2) +
  geom_hline(yintercept = 4, linetype = "dashed") +
  # Alert level 2 (NOAA)*
  geom_hline(yintercept = 8, linetype = "dashed") +
  annotate("rect", xmin = as.Date("2005-01-01"), xmax = as.Date("2021-01-01"), 
           ymin = 8, ymax = Inf, fill = "#d64541", alpha = .2) +
  # DHW
  geom_path(col = "black") +
  # DHW events
  geom_vline(xintercept = as.Date("2010-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2016-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2019-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2020-06-12"), col = "red") +
  # Labs and lims
  labs(x = NULL, y = "DHW (째C)") +
  lims(x = c(as.Date("2005-01-01"), as.Date("2021-01-01")), y = c(0, 10)) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# 1.3 Bleaching prevalence --

plot_blch <- data_bleaching %>% 
  group_by(date) %>% 
  summarise(bleaching_mean = mean(bleaching),
            bleaching_sd = sd(bleaching)) %>% 
  ggplot() +
  geom_pointrange(aes(x = date, y = bleaching_mean,
                      ymin = bleaching_mean-bleaching_sd, 
                      ymax = bleaching_mean+bleaching_sd)) +
  geom_path(aes(x = date, y = bleaching_mean)) +
  # DHW events
  geom_vline(xintercept = as.Date("2010-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2016-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2019-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2020-06-12"), col = "red") +
  # Labs and lims
  labs(x = NULL, y = "Bleaching (%)") +
  lims(x = c(as.Date("2005-01-01"), as.Date("2021-01-01"))) +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

# 1.4 HCC --

plot_hcc <- data_benthic %>%
  filter(category == "Hard corals") %>% 
  group_by(replicate, site, island, date) %>% 
  summarise(cover = sum(cover)) %>% 
  ungroup() %>% 
  group_by(date) %>% 
  summarise(cover_mean = mean(cover),
            cover_sd = sd(cover)) %>% 
  ungroup() %>% 
  ggplot() +
  geom_pointrange(aes(x = date, y = cover_mean,
                      ymin = cover_mean-cover_sd, 
                      ymax = cover_mean+cover_sd)) +
  geom_path(aes(x = date, y = cover_mean)) +
  # DHW events
  geom_vline(xintercept = as.Date("2010-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2016-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2019-06-12"), col = "red") +
  geom_vline(xintercept = as.Date("2020-06-12"), col = "red") +
  # Labs and lims
  labs(x = NULL, y = "HCC (%)") +
  lims(x = c(as.Date("2005-01-01"), as.Date("2021-01-01")))

# 2. Combine plots ----

plot_grid(plot_sst, plot_dhw, plot_blch, plot_hcc, nrow = 4, align = "v", labels = c("A", "B", "C", "D"))

# 3. Save the plot ----

ggsave("./../figs/04_combined-plots-sst-dhw-blch-hcc.png", height = 9, width = 5) # PNG

# 4. Remove useless objects ----

rm(plot_sst, plot_dhw, plot_blch, plot_hcc)

```

## Figure 2

### Sub-figure 1

```{r}

# 1. Transform the data for the NMDS ----

data_nmds <- data_benthic %>% 
  # Sum by replicate (remove subcategory)
  group_by(replicate, site, island, year, category) %>% 
  summarise(cover = sum(cover)) %>% 
  ungroup() %>% 
  # Mean by year and category
  group_by(year, category) %>% 
  summarise(cover = mean(cover)) %>% 
  ungroup() %>% 
  # Sum for each island and year, and standardization to 100
  group_by(year) %>% 
  mutate(total = sum(cover)) %>% 
  ungroup() %>% 
  mutate(cover = (cover*100)/total) %>% 
  # Misc.
  select(-total) %>% 
  pivot_wider(names_from = category, values_from = cover) %>% 
  replace(., is.na(.), 0) %>% # Replace all NA by 0
  column_to_rownames(var = "year") # Put year as rowname for the NMDS
  
# 2. Make the NMDS ----

data_nmds_transf <- decostand(data_nmds, "chi.square")

nmds_results <- metaMDS(data_nmds_transf, distance = "bray", k = 2, trymax = 100)

# 3. Extract coordinates ----

# 3.1 Extract year coordinates --

nmds_results_year <- data.frame(year = rownames(nmds_results$points),
                                NMDS1 = nmds_results$points[,1],
                                NMDS2 = nmds_results$points[,2]) %>% 
  left_join(., data_nmds %>% select("Hard corals") %>% rownames_to_column(var = "year")) %>% 
  rename(HCC = "Hard corals")

# 3.2 Extract category coordinates --

nmds_results_cat <- data.frame(category = rownames(nmds_results$species), 
                               NMDS1 = nmds_results$species[,1], 
                               NMDS2 = nmds_results$species[,2])

# 4. Make the plot ----

nmds_subfig1 <- ggplot() +
  # category data
  geom_segment(data = nmds_results_cat, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), color = "lightgrey") +
  geom_text_repel(data = nmds_results_cat, 
                  aes(x = NMDS1, y = NMDS2, label = category), 
                  family = font_choose_graph, size = 3, color = "darkgrey") +
  # Year data
  geom_path(data = nmds_results_year, 
             aes(x = NMDS1, y = NMDS2), size = 0.6,
             color = "black") +
  geom_point(data = nmds_results_year, 
             aes(x = NMDS1, y = NMDS2, fill = HCC),
             color = "black", shape = 21, size = 2) +
  scale_fill_continuous(low = "#4183d7", high = "#f22613") +
  geom_text_repel(data = nmds_results_year, 
                  aes(x = NMDS1, y = NMDS2, label = year), 
                  size = 3, family = font_choose_graph,
                  color = "black") +
  # 2D stress
  annotate(geom = "text", Inf, -Inf, 
           label = paste0("2D stress = ", round(nmds_results$stress, 3)), 
           color = "black",
           hjust = 1.1, vjust = -2,
           family = font_choose_graph) +
  # Bleaching events
  annotate(geom = "point", 
           x = ((nmds_results_year["2015","NMDS1"] + nmds_results_year["2016","NMDS1"])/2),
           y = ((nmds_results_year["2015","NMDS2"] + nmds_results_year["2016","NMDS2"])/2),
           shape = 4, color = "red") +
  annotate(geom = "point", 
           x = ((nmds_results_year["2009","NMDS1"] + nmds_results_year["2010","NMDS1"])/2),
           y = ((nmds_results_year["2009","NMDS2"] + nmds_results_year["2010","NMDS2"])/2),
           shape = 4, color = "red") +
  # Misc.
  labs(x = "nMDS 1", y = "nMDS 2", title = "A. Benthic community") +
  theme(panel.grid = element_blank())

```

### Sub-figure 2

```{r}

data_nmds <- data_benthic %>% 
  filter(category == "Hard corals") %>% 
  # Sum by replicate
  group_by(replicate, site, island, year, subcategory) %>% 
  summarise(cover = sum(cover)) %>% 
  ungroup() %>% 
  # Mean by year and subcategory
  group_by(year, subcategory) %>% 
  summarise(cover = mean(cover)) %>% 
  ungroup() %>% 
  # Sum for each island and year, and standardization to 100
  group_by(year) %>% 
  mutate(total = sum(cover)) %>% 
  ungroup() %>% 
  mutate(cover = (cover*100)/total) %>% 
  # Misc.
  select(-total) %>% 
  filter(year != 2005, year != 2007) %>% 
  pivot_wider(names_from = subcategory, values_from = cover) %>% 
  replace(., is.na(.), 0) %>% # Replace all NA by 0
  column_to_rownames(var = "year") # Put year as rowname for the NMDS

# 2. Make the NMDS ----

data_nmds_transf <- decostand(data_nmds, "chi.square")

nmds_results <- metaMDS(data_nmds_transf, distance = "bray", k = 2, trymax = 100)

# 3. Plot the results ----

# 3.1 Extract year coordinates --

nmds_results_year <- data.frame(year = rownames(nmds_results$points),
                                NMDS1 = nmds_results$points[,1],
                                NMDS2 = nmds_results$points[,2])

# 3.2 Extract category coordinates --

nmds_results_cat <- data.frame(category = rownames(nmds_results$species), 
                               NMDS1 = nmds_results$species[,1], 
                               NMDS2 = nmds_results$species[,2])

# 3.3 Make the plot --

ggplot() +
  # category data
  geom_segment(data = nmds_results_cat, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), color = "lightgrey") +
  geom_text_repel(data = nmds_results_cat, 
            aes(x = NMDS1, y = NMDS2, label = category), 
            family = font_choose_graph, size = 3, color = "darkgrey") +
  theme(panel.grid = element_blank()) +
  # Year data
  geom_path(data = nmds_results_year, 
             aes(x = NMDS1, y = NMDS2), size = 0.6,
             color = "black") +
  geom_point(data = nmds_results_year, 
             aes(x = NMDS1, y = NMDS2),
             color = "black", shape = 21, size = 2) +
  geom_text_repel(data = nmds_results_year, 
                  aes(x = NMDS1, y = NMDS2, label = year), 
                  size = 3, family = font_choose_graph,
                  color = "black") +
  # 2D stress
  annotate(geom = "text", Inf, -Inf, 
           label = paste0("2D stress = ", round(nmds_results$stress, 3)), 
           color = "black",
           hjust = 1.1, vjust = -2,
           family = font_choose_graph) +
  # Misc.
  labs(x = "nMDS 1", y = "nMDS 2", title = "B. Hard coral shape")

```

### Sub-figure 3

```{r}

# 1. Preliminary transformation of the data for the NMDS ----

nmds_results <- data_recruits %>% 
  # Mean by island, year and category
  group_by(year, genus) %>% 
  summarise(recruit = mean(recruit)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = genus, values_from = recruit) %>% 
  replace(., is.na(.), 0) %>% # Replace all NA by 0
  column_to_rownames(var = "year") %>% # Put year as rowname for the NMDS
  metaMDS(., distance = "bray", k = 2, trymax = 100) # NMDS

# 2. Plot the results ----

# 2.1 Extract year coordinates --

nmds_results_year <- data.frame(year = rownames(nmds_results$points),
                                NMDS1 = nmds_results$points[,1],
                                NMDS2 = nmds_results$points[,2])

# 2.2 Extract category coordinates --

nmds_results_recruit <- data.frame(category = rownames(nmds_results$species), 
                                   NMDS1 = nmds_results$species[,1], 
                                   NMDS2 = nmds_results$species[,2])

# 2.3 Make the plot --

ggplot() +
  # category data
  
    geom_segment(data = nmds_results_recruit, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), color = "lightgrey") +
   geom_text_repel(data = nmds_results_recruit, 
            aes(x = NMDS1, y = NMDS2, label = category), 
            family = font_choose_graph, size = 3, color = "darkgrey") +
  theme(panel.grid = element_blank()) +

  # Year data
  geom_path(data = nmds_results_year, 
            aes(x = NMDS1, y = NMDS2), size = 0.6,
            color = "black") +
  geom_point(data = nmds_results_year, 
             aes(x = NMDS1, y = NMDS2),
             color = "black", fill = "red", shape = 21, size = 2) +
  geom_text_repel(data = nmds_results_year, 
                  aes(x = NMDS1, y = NMDS2, label = year), 
                  size = 3, family = font_choose_graph,
                  color = "black") +
  # 2D stress
  annotate(geom = "text", Inf, -Inf, 
           label = paste0("2D stress = ", round(nmds_results$stress, 3)), 
           color = "black",
           hjust = 1.1, vjust = -2,
           family = font_choose_graph) +
  # Misc.
  labs(x = "NMDS 1", y = "NMDS 2", title = "B. Coral recruits")

```

### Assemble sub-figures

```{r fig.height=5, fig.width=15}

# 1. Combine plots ----

plot_grid(nmds_subfig1, nmds_subfig1, nmds_subfig1, nrow = 1, align = "h")

# 2. Save the plot ----

ggsave("./../figs/04_combined-plots-sst-dhw-blch-hcc.png", height = 9, width = 5) # PNG

# 3. Remove useless objects ----

rm(plot_sst, plot_dhw, plot_blch, plot_hcc)

```

# Supplementary figures

## Supp. figure 1

```{r fig.height=10, fig.width=25}

# 1. Make the plot ----

data_benthic %>% 
  # Sum by replicate (remove subcategory)
  group_by(replicate, site, island, date, category) %>% 
  summarise(cover = sum(cover)) %>% 
  ungroup() %>% 
  # Mean by island, date and category
  group_by(island, date, category) %>% 
  summarise(cover = mean(cover)) %>% 
  ungroup() %>% 
  # Sum for each island and date, and standardization to 100
  group_by(island, date) %>% 
  mutate(total = sum(cover)) %>% 
  ungroup() %>% 
  mutate(cover = (cover*100)/total) %>% 
  # Plot 
  ggplot(data = ., aes(x = date, y = cover, fill = category)) +
    geom_vline(xintercept = as.Date("2015-12-05"), color = "red", size = 1.5) +
    geom_vline(xintercept = as.Date("2009-12-05"), color = "red", size = 1.5) +
    geom_bar(stat = "identity", color = "black", size = 0.25) +
    scale_fill_manual(values = palette_category) +
    facet_wrap(~island, ncol = 7) +
    labs(x = "Year", y = "Benthic cover (%)", fill = "Category")

# 2. Save the plot ----

ggsave("./../figs/supp-fig-1_benthic-composition-by-island.png", height = 10, width = 25) # PNG

```

## Supp. figure 2

```{r fig.height=10, fig.width=25}

# 1. Make the plot ----

data_benthic %>% 
  # Filter HCC
  filter(category == "Hard corals") %>% 
  # Extract the HCC shape
  mutate(subcategory = str_split_fixed(subcategory, " ", 2)[,2]) %>% 
  # Sum by replicate
  group_by(replicate, site, island, date, subcategory) %>% 
  summarise(cover = sum(cover)) %>% 
  ungroup() %>% 
  # Mean by island, date and subcategory
  group_by(island, date, subcategory) %>% 
  summarise(cover = mean(cover)) %>% 
  ungroup() %>% 
  # Plot 
  ggplot(data = ., aes(x = date, y = cover, fill = subcategory)) +
    geom_vline(xintercept = as.Date("2015-12-05"), color = "red", size = 1.5) +
    geom_vline(xintercept = as.Date("2009-12-05"), color = "red", size = 1.5) +
    geom_bar(stat = "identity", color = "black", size = 0.25) +
    facet_wrap(~island, ncol = 7) +
    labs(x = "Year", y = "Benthic cover (%)", fill = "Coral shape")

# 2. Save the plot ----

ggsave("./../figs/supp-fig-2_hard-coral-shape-by-island.png", height = 10, width = 25) # PNG

```

## Supp. figure 3

```{r fig.height=10, fig.width=25}

# 1. Make the plot ----

ggplot(data_recruits, aes(x = year, y = recruit, fill = genus)) +
  geom_bar(stat = "identity", color = "black", size = 0.25) +
  facet_wrap(~island, ncol = 7) +
  labs(x = "Year", y = "Recruits (n.m2)", fill = "Genus")

# 2. Save the plot ----

ggsave("./../figs/supp-fig-3_number-recruits-by-genus-and-island.png", height = 10, width = 25) # PNG

```

## Supp. figure 4

```{r fig.height=8, fig.width=12}

# 1. Transform the data for the NMDS ----

data_nmds <- data_benthic %>% 
  # Sum by replicate (remove subcategory)
  group_by(replicate, site, island, year, category) %>% 
  summarise(cover = sum(cover)) %>% 
  ungroup() %>% 
  # Mean by year and category
  group_by(island, year, category) %>% 
  summarise(cover = mean(cover)) %>% 
  ungroup() %>% 
  # Sum for each island and year, and standardization to 100
  group_by(island, year) %>% 
  mutate(total = sum(cover)) %>% 
  ungroup() %>% 
  mutate(cover = (cover*100)/total) %>% 
  # Misc.
  mutate(label = paste(island, year, sep = "_")) %>% 
  select(-total, -island, -year) %>% 
  pivot_wider(names_from = label, values_from = cover) %>% 
  replace(., is.na(.), 0) %>% # Replace all NA by 0
  column_to_rownames(var = "category") # Put year as rowname for the NMDS

# 2. Make the NMDS ----

data_nmds_transf <- decostand(data_nmds, "chi.square")

nmds_results <- metaMDS(data_nmds_transf, distance = "bray", k = 2, trymax = 100)

# 3. Extract coordinates ----

# 3.1 Extract year coordinates --

nmds_results_cat <- data.frame(category = rownames(nmds_results$points),
                                NMDS1 = nmds_results$points[,1],
                                NMDS2 = nmds_results$points[,2])

# 3.2 Extract category coordinates --

nmds_results_year <- data.frame(label = rownames(nmds_results$species), 
                               NMDS1 = nmds_results$species[,1], 
                               NMDS2 = nmds_results$species[,2]) %>% 
  mutate(island = str_split_fixed(label, "_", 2)[,1],
         year = str_split_fixed(label, "_", 2)[,2])

# 3.3 Create the convex hull --

nmds_hull_year <- nmds_results_year %>%
  group_by(year) %>% 
  slice(chull(NMDS1, NMDS2))

# 4. Make the plot ----

ggplot() +
  # Year data
  geom_segment(data = nmds_results_cat, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), color = "lightgrey") +
  geom_text_repel(data = nmds_results_cat, 
                  aes(x = NMDS1, y = NMDS2, label = category), 
                  size = 3, family = font_choose_graph,
                  color = "darkgrey") +
  # category data
  geom_point(data = nmds_results_year, aes(x = NMDS1, y = NMDS2, fill = year), 
             shape = 21, color = "black", show.legend = FALSE) +
  geom_polygon(data = nmds_hull_year, aes(x = NMDS1, y = NMDS2, fill = year), 
               alpha = 0.3, show.legend = FALSE, color = "grey", linetype = "dashed") +
  # Misc.
  labs(x = "NMDS 1", y = "NMDS 2") +
  theme(panel.grid = element_blank()) +
  facet_wrap(~year, nrow = 3)

# 5. Save the plot ----

ggsave("./../figs/supp-fig-4_nmds-benthic-composition-by-year-and-island.png", height = 8, width = 12) # PNG

```

## Supp. figure 5

```{r fig.height=8, fig.width=12}

# 1. Transform the data for the NMDS ----

data_nmds <- data_benthic %>% 
  filter(category == "Hard corals") %>% 
  # Sum by replicate
  group_by(replicate, site, island, year, subcategory) %>% 
  summarise(cover = sum(cover)) %>% 
  ungroup() %>% 
  # Mean by year and subcategory
  group_by(island, year, subcategory) %>% 
  summarise(cover = mean(cover)) %>% 
  ungroup() %>% 
  # Sum for each island and year, and standardization to 100
  group_by(island, year) %>% 
  mutate(total = sum(cover)) %>% 
  ungroup() %>% 
  mutate(cover = (cover*100)/total) %>% 
  # Misc.
  mutate(label = paste(island, year, sep = "_")) %>% 
  select(-total, -island, -year) %>% 
  pivot_wider(names_from = label, values_from = cover) %>% 
  replace(., is.na(.), 0) %>% # Replace all NA by 0
  column_to_rownames(var = "subcategory") # Put year as rowname for the NMDS

# 2. Make the NMDS ----

data_nmds_transf <- decostand(data_nmds, "chi.square")

nmds_results <- metaMDS(data_nmds_transf, distance = "bray", k = 2, trymax = 100)

# 3. Plot the results ----

# 3.1 Extract year coordinates --

nmds_results_year <- data.frame(label = rownames(nmds_results$species), 
                               NMDS1 = nmds_results$species[,1], 
                               NMDS2 = nmds_results$species[,2]) %>% 
  mutate(island = str_split_fixed(label, "_", 2)[,1],
         year = str_split_fixed(label, "_", 2)[,2])

# 3.2 Extract category coordinates --

nmds_results_cat <- data.frame(subcategory = rownames(nmds_results$points),
                                NMDS1 = nmds_results$points[,1],
                                NMDS2 = nmds_results$points[,2])

# 3.3 Create the convex hull --

nmds_hull_year <- nmds_results_year %>%
  group_by(year) %>% 
  slice(chull(NMDS1, NMDS2))

# 4. Make the plot ----

ggplot() +
  # Category data
  geom_segment(data = nmds_results_cat, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), color = "lightgrey") +
  geom_text_repel(data = nmds_results_cat, 
                  aes(x = NMDS1, y = NMDS2, label = subcategory), 
                  size = 3, family = font_choose_graph,
                  color = "darkgrey") +
  # Year data
  geom_point(data = nmds_results_year, aes(x = NMDS1, y = NMDS2, fill = year), 
             shape = 21, color = "black", show.legend = FALSE) +
  geom_polygon(data = nmds_hull_year, aes(x = NMDS1, y = NMDS2, fill = year), 
               alpha = 0.3, show.legend = FALSE, color = "grey", linetype = "dashed") +
  # Misc.
  labs(x = "NMDS 1", y = "NMDS 2") +
  theme(panel.grid = element_blank()) +
  facet_wrap(~year, nrow = 3)

# 5. Save the plot ----

ggsave("./../figs/supp-fig-5_nmds-coral-shape-by-year-and-island.png", height = 8, width = 12) # PNG

```

## Supp. figure 6

```{r fig.height=8, fig.width=12}

# 1. Transform the data for the NMDS ----

data_nmds <- data_recruits %>% 
  filter(!(genus %in% c("Echinopora sp.", "Platygyra sp."))) %>% 
  mutate(label = paste(island, year, sep = "_")) %>% 
  select(-island, -year) %>% 
  pivot_wider(names_from = label, values_from = recruit) %>% 
  replace(., is.na(.), 0) %>% # Replace all NA by 0
  column_to_rownames(var = "genus")

# 2. Make the NMDS ----

data_nmds_transf <- decostand(data_nmds, "chi.square")

nmds_results <- metaMDS(data_nmds_transf, distance = "bray", k = 2, trymax = 100)

# 3. Plot the results ----

# 3.1 Extract year coordinates --

nmds_results_year <- data.frame(label = rownames(nmds_results$species), 
                                NMDS1 = nmds_results$species[,1], 
                                NMDS2 = nmds_results$species[,2]) %>% 
  mutate(island = str_split_fixed(label, "_", 2)[,1],
         year = str_split_fixed(label, "_", 2)[,2])

# 3.2 Extract category coordinates --

nmds_results_cat <- data.frame(genus = rownames(nmds_results$points),
                               NMDS1 = nmds_results$points[,1],
                               NMDS2 = nmds_results$points[,2])

# 3.3 Create the convex hull --

nmds_hull_year <- nmds_results_year %>%
  group_by(year) %>% 
  slice(chull(NMDS1, NMDS2))

# 4. Make the plot ----

ggplot() +
  # Category data
  geom_segment(data = nmds_results_cat, aes(x = 0, y = 0, xend = NMDS1, yend = NMDS2), color = "lightgrey") +
  geom_text_repel(data = nmds_results_cat, 
                  aes(x = NMDS1, y = NMDS2, label = genus), 
                  size = 3, family = font_choose_graph,
                  color = "darkgrey") +
  # Year data
  geom_point(data = nmds_results_year, aes(x = NMDS1, y = NMDS2, fill = year), 
             shape = 21, color = "black", show.legend = FALSE) +
  geom_polygon(data = nmds_hull_year, aes(x = NMDS1, y = NMDS2, fill = year), 
               alpha = 0.3, show.legend = FALSE, color = "grey", linetype = "dashed") +
  # Misc.
  labs(x = "NMDS 1", y = "NMDS 2") +
  theme(panel.grid = element_blank()) +
  facet_wrap(~year, nrow = 3)

# 5. Save the plot ----

ggsave("./../figs/supp-fig-6_nmds-coral-recruits-by-year-and-island.png", height = 8, width = 12) # PNG

```

# Reproducibility

```{r reprod}

# 1. Reproducibility ----

sessionInfo()

```

---
Jeremy WICQUART | jeremywicquart@gmail.com | `r format(Sys.time())`